---
title: "Practical.Machine.Learning.Course.Project"
author: "Coursera_Fan"
date: "June 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
```

### Human Activity Recognition (HAR) research focusses on predciting which activity was performed at specific time.
*  In this project we are trying to predict if certain activity was performed correctly or not, based upon data provided.
    1.  objective of this study is to investigate, how well an activity (excercise in this case) was performed. This can have applications in sports training.
    2.  Six males age 20-28 performed dumbell biceps curl in five different ways, named A,B,C,D,E, were A is correct way to do excercise and B,C,D,E are incorrect ways.


```{r cars, echo=T, warning=FALSE}
library(caret); library(randomForest)
```

#### download the training & test dataset
```{r, echo=T}
har.training.dwnld <- download.file(url= "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile= "HARtraining.csv")
har.test.dwnld <- download.file(url= "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv" , destfile= "HARtesting.csv")
```
####  read the training as har. Read the testing dataset as harTestSet
```{r, echo=T, results=T}
hartraining <- read.csv("HARtraining.csv", sep = ",", header = T, na.strings = c("NA",".","#DIV/0!"))
dim(hartraining)

harTestSet <- read.csv("HARtesting.csv", sep = ",", header = T, na.strings = c("NA",".","#DIV/0!"))
```
####  removing the missing vals from hartraining & HarTestSet
*  This step will reduce the number of cols as well to 60 from 160
```{r, echo=T}
hartraining <- hartraining[,colSums(is.na(hartraining))==0]
harTestSet <- harTestSet[, which(colSums(is.na(harTestSet))==0)]

```
####  selecting columns with numeric class only
```{r, echo=T, results=T}
numeric.Cols <- which(sapply(hartraining, class) == "numeric")
hartrain <- hartraining[,numeric.Cols]
hartrain$classe <- hartraining$classe
dim(hartrain)
```
#### partioning hartrain into training & validation
```{r, echo=T}
inTrain <- createDataPartition(y = hartrain$class, p = 3/4, list=F)
train <- hartrain[inTrain,]
validation <-hartrain[-inTrain,]
```
#### showing the dimensions of all datasets created
```{r, echo=F, results=T, warning=F}
print(paste0("hartraining rows-->", nrow(hartraining),", ", "hartraining cols -->", ncol(hartraining)))
print(paste0("hartrain rows-->", nrow(hartrain),", ", "hartrain cols -->", ncol(hartrain)))
print(paste0("train rows-->", nrow(train),", ", "train cols -->", ncol(train)))
print(paste0("validation rows-->", nrow(validation),", ", "validation cols -->", ncol(validation)))
print(paste0("harTestSet rows-->", nrow(harTestSet),", ", "harTestSet cols -->", ncol(harTestSet)))
```
#### Creating a model for train dataset
```{r, echo=T}
modTrain <- train(classe~. , preProcess=c("knnImpute", "center", "scale"), method="rf", data=train)
```
#### Applying this model to create predict outcomes for validation set
```{r, echo=T}
predictedVal <- predict(modTrain,validation)
```
#### checking the accuracy of our model prediction against validation data
```{r, echo=T, results=F,warning=F}
accuracy  <- confusionMatrix(predictedVal,validation$classe)
```
```{r, echo=T}
print(accuracy$overall)
```
####  out of sample error rate
```{r, echo=T}
in.sample.error.rate <- (1)-(accuracy$overall)[[1]]
print(paste0("in sample error rate-->",in.sample.error.rate)*100, " %")
outOfSampleAccuracy <- sum(predictedVal == validation$classe)/length(predictedVal)
out.of.sample.error <- (1-outOfSampleAccuracy)
print(paste0("out of sample error -->", (out.of.sample.error)*100, " %"))
```

####  Outcomes for harTestSet using our predition model
```{r, echo=T, results=F}
quizPredict<- predict(modTrain,harTestSet)
```
#### Here are the final results from prediction model for the quiz
```{r, echo=F, results=T}
as.data.frame(quizPredict)
```